---
title: "Homework 6"
author: "Shritama Ray"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(broom)
library(p8105.datasets)
library(viridis)
library(reshape2)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1: Central Park Weather

```{r import data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

### Problem 2: Washington Post Homocide Data

**Read in & tidy the data:**
```{r read and tidy data}
#Read in data from github
homicides = read.csv('https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv') %>%
  janitor::clean_names()%>%
  mutate(
    city_state = str_c(city, ", ", state),
    solved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age))%>%
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"), victim_race %in% c("White", "Black"))
```

**Logistic Regression for Baltimore, MD**
```{r baltimore}
baltimore_df = 
  homicides %>% 
  filter(city_state == "Baltimore, MD") %>%
  select(solved, victim_age, victim_race, victim_sex) 

baltimore_logistic =
  glm(solved ~ victim_age + victim_race + victim_sex, data = baltimore_df, family = binomial()) %>%
  broom::tidy()%>%
  filter(str_detect(term, "Male"))%>%
  select(term, estimate, std.error)
```

**GLM for all Cities**
```{r all cities}
glm_all =
  homicides %>%
  select(city_state, solved, victim_age, victim_race, victim_sex) %>%
  nest(data = -city_state)%>%
  mutate(
    models = map(data, ~glm(solved ~ victim_age + victim_race + victim_sex, 
    na.action = na.omit, data = .x)),
    results = map(models, broom::tidy)) %>%
  select(-data, -models) %>%
  unnest(results) %>%
  filter(str_detect(term, "Male"))%>%
  mutate(lower_95 = estimate - 1.96*std.error, upper_95 = estimate + 1.96*std.error) %>%
  select(city_state, term, estimate, lower_95, upper_95) 

#Plot
glm_all %>%
  mutate(fct_reorder(city_state, estimate))%>%
  ggplot(aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_point() +
geom_errorbar(aes(ymin = lower_95, ymax = upper_95)) +
  labs(
    title = "Odds Ratio of Solved Homicides for Male vs. Female Victims by City",
    x = "City, State",
    y = "Odds Ratio"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.7, hjust = 1),
    plot.title = element_text(hjust = 0.5))
```


### Problem 3: Birthweight Regression

Load & clean data:
```{r birthweight load}
birthweight = read_csv("./birthweight.csv") %>%
  janitor::clean_names() %>%
  mutate(
    #convert from numeric to factor variables
    babysex = factor(babysex),
    frace = factor(frace),
    malform = factor(malform),
    mrace = factor(mrace))

str(birthweight)

```

Propose a regression model:
```{r birthweight stepwise}
#Fit a model using backwards step-wise selection (data-driven)
lm_backward <- lm(bwt ~., data = birthweight)
step(lm_backward, direction = "backward") 
```

I created a data-driven model using backwards stepwise selection. This process starts with all variables as covariates to predict the outcome (birthweight). It then one by one, removes the lease significant variables from the model until a final model is fitted with only the most significant covariates remaining.

The final model created through this process is:
bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + mheight + mrace + parity + ppwt + smoken

Plot of model residuals against fitted values:
```{r backwards residuals}
birthweight %>%
  add_residuals(lm_backward) %>%
  add_predictions(lm_backward) %>%
  ggplot(aes(x = pred, y = resid)) + geom_point() +
  labs(title = "Backwards Selection Model Residuals vs. Fitted Values")
```

Create 2 other models:
```{r comparison models}
#Model length at birth and gestation age as predictors
model_a = lm(bwt ~ blength + gaweeks, data = birthweight)

#Head circumference, length, sexm and all interactions
model_b = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = birthweight)
```

Compare the 3 models using cross-validated prediction error:
```{r model comparison}
#Obtain RMSEs:
cv_df = crossv_mc(birthweight, 100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble),
    backwards = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + mheight + mrace + parity + ppwt + smoken, data = .x)),
    model_a = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model_b = map(train, ~lm(bwt ~ bhead + blength + babysex 
                                     + bhead*blength 
                                     + bhead*babysex 
                                     + blength*babysex 
                                     + bhead*blength*babysex, data = .x)), 
    rmse_backwards = map2_dbl(backwards, test, ~rmse(model = .x, data = .y)),
    rmse_model_a = map2_dbl(model_a, test, ~rmse(model = .x, data = .y)),
    rmse_model_b = map2_dbl(model_b, test, ~rmse(model = .x, data = .y)))

#Plot prediction error distribution for each candidate model:
cv_df %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(title = "RMSEs for 3 Candidate Models", x = "Model", y = "RMSE")
```
